---
title: "PDST"
author: "Sam Newman"
date: "2023-10-09"
output: html_document
---

Background:
The data comes from one qualitative question asked in a survey that was given to
a universities (named disclosed for confidentiality) 2nd year MPP students (n=79)
from 2022-2023. The question was free-response and asked students for their opinion on 
increasing Professional Degree Supplemental Tuition (PDST) fees for MPP students. 
The fees act like a tax that all MPP students pay and the money is then distributed
to other students in great financial need. The data is partially cleaned in the script, 
but mostly cleaned in word. The results are then visualized in a bar chart. 

Outline: 
1. Clean Data
2. Visualize Data

# load libraries 
```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(stringr)
library(gt)
```
#set working directory 
```{r}
setwd("C:/Grad School/Student Affairs/Exit Survey")
```


# load clean exit survey data 2019-2023 (object is created in 'Exit Survey_Analysis_2')
```{r}
sub_e2019_2023 <- readRDS("sub_e2019_2023.rds")

```

Since pdst is mostly qualitative I am going to export the data into an excel sheet. But first, I will clean the data by consolidating NAs and then removing them form the data set. 

# create subset for data and intial investigation variable 
```{r}
pdst <- sub_e2019_2023 %>%  select(pdst, year)

#160 NAs to start 
sum(is.na(pdst$pdst))

#269 observations
length(pdst$pdst)

view(pdst)

#Types of NA: c('n/a', N/a, N/A, N/A, NA)
#convert the NAs above into NA that R recognizes 
pdst_clean <- pdst %>%  filter(
  pdst != 'n/a', 
  pdst != 'N/a', 
  pdst != 'N/A', 
  pdst != 'NA', 
  pdst != 'yes', 
  pdst != 'Unsure', 
  pdst != '-', 
  pdst != 'Na', 
  pdst != "No response", 
  pdst != "No comment", 
  pdst != "none")

#length is now 84 observatoins
length(pdst_clean$pdst)

#0 NAs
sum(is.na(pdst_clean))

#save file as excel sheet to perform qualitative analysis 
write_xlsx(pdst_clean, "pdst_clean.xlsx")

```

I have removed 3 comments from the observations.
I exported the file to Excel where I read each entry and assigned it to "Yes", "No", or "Other". I then tallied the results in Word. 

*Create Bar Chart* 
I will create a df that contains the tallies for the responses and then visualize the results in a bar chart. 
                          
# create data frame containing count for PDST
```{r}
pdst_count <- data.frame(
 response = c("Yes", "No", "Other"),
 mentions = as.numeric(c(21, 41, 16)))


#check class types 
class(pdst_count$response) #character

class(pdst_count$mentions) #numeric 

```
# factor answers 
```{r}
#create new object 
factor_pdst_count <- pdst_count

factor_pdst_count$response <- factor(
 factor_pdst_count$response, levels = c(
             "Yes", 
              "No",
              "Other"))

#check class type (it's factor - good)
class(factor_pdst_count$response)

```
#create bar chart for favorite courses
```{r}
ggplot(factor_pdst_count, aes(x = response , 
                             y = mentions, 
                             )) +
  geom_bar(
         stat = 'identity',
         fill = "lightblue",
           position = "dodge",  #makes bar chart side by side
           color = "black", #adds outline color to bins
         na.rm = TRUE, 
         width = .5) +
  scale_y_continuous(
    limits = c(0, 45),                
    breaks = seq(0,45, 5)) +
          labs (x = "", 
                y = "Number of Student Responses",
                title = "Student Responses to Increasing PDST Fees \n (2022-2023)")+
       theme(
         axis.text = element_text( hjust = .5, size =10.5), 
         plot.caption = element_text(hjust = 1, ), 
         plot.title = element_text(hjust = .5, face = "bold", size = 12))+
          labs(caption ="n= 78 student responses") 
```
